---
title: "GAM Report - Satisfaction Level"
output:
  html_document:
    toc: true
    number_sections: false
    df_print: paged
    css: styles.css
params:
  target: NULL
  gam_model: NULL
  numeric_plots: NULL
  categorical_plots: NULL
  residuals_plot: NULL
  obs_vs_pred_plot: NULL
  smooth_plots: NULL
  appraise_plot: NULL
  rmse: NULL
  ref_levels_df: NULL
  gvif_tbl: NULL
  cv_results: NULL
  cv_summary: NULL
  categorical_features: NULL
  importance_tbl: NULL
---

```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(knitr)
library(kableExtra)
library(gratia)
library(mgcv)
library(performance)
library(dplyr)
library(tidyr)
library(tibble)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE
)
```

# Model Performance Metrics

- Number of sessions: `r nobs(params$gam_model)`
- Adjusted R²: `r round(summary(params$gam_model)$r.sq, 3)`
- Deviance explained: `r round(summary(params$gam_model)$dev.expl[1], 3)`
- AIC: `r format(round(AIC(params$gam_model), 2), scientific = FALSE)`
- RMSE: `r round(params$rmse, 2)`

**Cross-Validated Performance (K-fold CV)**

```{r}
if (!is.null(params$cv_summary)) {

  kable(
    params$cv_summary,
    digits = 3
  ) %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "condensed")
    )

} else {
  cat("Cross-validation results not provided.")
}
```

```{r}
if (!is.null(params$cv_results)) {

  # Transpose the table
  cv_results_t <- as.data.frame(t(params$cv_results))
  cv_results_t <- cv_results_t[-1, ]

  # Optionally rename the columns to preserve fold numbers
  colnames(cv_results_t) <- paste0("Fold ", seq_len(ncol(cv_results_t)))

  # Print transposed table with kable
  kable(
    cv_results_t,
    digits = 3,
    caption = "Fold-level RMSE and MAE from K-fold CV"
  ) %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "condensed")
    )

} else {
  cat("Fold-level CV results not provided.")
}
```

# Key Driver Ranking

Comparison of individual deviance (gam.hp) and per session contribution variance.

```{r}
if (!is.null(params$importance_tbl)) {

  driver_tbl <- params$importance_tbl %>%
    dplyr::select(
      covariate,
      Unique,
      Average.share,
      Individual,
      `I.perc(%)`,
      TermSD_pct,
      TermSD_unweighted_pct,
      TermSD,
      TermSD_unweighted
    ) %>%
    dplyr::mutate(
      Unique        = round(Unique, 3),
      Average.share = round(Average.share, 3),
      Individual    = round(Individual, 3),
      `I.perc(%)`   = round(`I.perc(%)`, 2),
      TermSD_pct   = round(TermSD_pct, 2),
      TermSD_unweighted_pct = round(TermSD_unweighted_pct, 2),
      TermSD        = round(TermSD, 2),
      TermSD_unweighted = round(TermSD_unweighted, 2)
    )

  kable(
    driver_tbl,
    caption = paste(
      "Key drivers sorted by gam.hp output (Perecentage individual contribution - 'I.perc(%)').",
      "Unique, Average.share, Individual, I.perc(%): Columns from gam.hp hierarchical partitioning.",
      "TermSD_pct: Relative Share of Standard Deviation of term contributions.",
      "TermSD: Standard Deviation of term contributions on the response scale (√Var)."
    )
  ) %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "condensed")
    )

} else {
  cat("Key driver importance not available.")
}
```

# Reference Levels for Categorical Features

```{r}
kable(params$ref_levels_df) %>%
  kable_styling(
    full_width = FALSE,
    position = "left",
    bootstrap_options = c("striped", "condensed")
  )
```

# Model Specification

```{r}
formula_str <- attr(params$gam_model, "formula_str")
```
**Full GAM formula:**
`r formula_str`

## GAM Model Summary - Target: `r params$target`


```{r}
sm <- summary(params$gam_model)
pt <- as.data.frame(sm$p.table)
pnames <- names(pt)

is_t <- "Pr(>|t|)" %in% pnames
is_z <- "Pr(>|z|)" %in% pnames

if (is_t) {

  param_tab <- as.data.frame(sm$p.table) %>%
    tibble::rownames_to_column("Term") %>%
    mutate(
      Variable = case_when(
        Term == "(Intercept)" ~ "(Intercept)",
        grepl("^year\\d{4}$", Term) ~ "year",
        TRUE ~ sub("([A-Z].*)$", "", Term)
      ),
      Level = case_when(
        Term == "(Intercept)" ~ NA_character_,
        grepl("^year(\\d{4})$", Term) ~ sub("^year", "", Term),
        TRUE ~ sub("^[^A-Z]+", "", Term)
      ),
      Level = ifelse(Level == "", NA, Level),
      Signif = symnum(
        `Pr(>|t|)`,
        cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
        symbols   = c("***", "**", "*", ".", " ")
      ),
      Estimate   = round(Estimate, 2),
      `Std. Error` = round(`Std. Error`, 2),
      `t value`  = round(`t value`, 2),
      `Pr(>|t|)` = formatC(`Pr(>|t|)`, format = "e", digits = 2),
    ) %>%
    select(Variable, Level, Estimate, `Std. Error`, `t value`, `Pr(>|t|)`, Signif)

  param_tab <- param_tab %>%
    dplyr::rename(`p-value` = `Pr(>|t|)`)  

  kable(param_tab,
        caption = "Parametric effects from the GAM") %>%
    kable_styling(full_width = FALSE, position = "left",
                  bootstrap_options = c("striped", "condensed"))

} else if (is_z) {

  param_tab <- as.data.frame(sm$p.table) %>%
    tibble::rownames_to_column("Term") %>%
    mutate(
      Variable = case_when(
        Term == "(Intercept)" ~ "(Intercept)",
        grepl("^year\\d{4}$", Term) ~ "year",
        TRUE ~ sub("([A-Z].*)$", "", Term)
      ),
      Level = case_when(
        Term == "(Intercept)" ~ NA_character_,
        grepl("^year(\\d{4})$", Term) ~ sub("^year", "", Term),
        TRUE ~ sub("^[^A-Z]+", "", Term)
      ),
      Level = ifelse(Level == "", NA, Level),
      Signif = symnum(
        `Pr(>|z|)`,
        cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
        symbols   = c("***", "**", "*", ".", " ")
      ),
      Estimate   = round(Estimate, 2),
      `Std. Error` = round(`Std. Error`, 2),
      `z value`  = round(`z value`, 2),
      `Pr(>|z|)` = formatC(`Pr(>|z|)`, format = "e", digits = 2),
    ) %>%
    select(Variable, Level, Estimate, `Std. Error`, `z value`, `Pr(>|z|)`, Signif)

  param_tab <- param_tab %>%
    dplyr::rename(`p-value` = `Pr(>|z|)`)  

  kable(param_tab,
        caption = "Parametric effects from the GAM") %>%
    kable_styling(full_width = FALSE, position = "left",
                  bootstrap_options = c("striped", "condensed"))
}
```

```{r}
smooth_tab <- as.data.frame(sm$s.table) %>%
  mutate(
    Signif = symnum(
      `p-value`,
      cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
      symbols   = c("***", "**", "*", ".", " ")
    ),
    edf       = round(edf, 2),
    F         = round(F, 2),  # nolint
    `p-value` = formatC(`p-value`, format = "e", digits = 2),
  ) %>%
  select(edf, Ref.df, F, `p-value`, Signif)  # nolint

kable(smooth_tab, caption = "Approximate significance of smooth terms") %>%
  kable_styling(full_width = FALSE, position = "left",
                bootstrap_options = c("striped", "condensed"))
```
Significance codes:  `0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1`

# Smooth Effects

**Spline k-index Diagnostics**

Note: The p-values in the k-index table are different from the smooth-term significance p-values.  
Higher p-values are desirable because they indicate there is no statistical evidence that increasing `k` would improve the fit.  
Low p-values suggest `k` may be too small, meaning the smooth might be underfitting.

```{r}
k_diag <- mgcv::k.check(params$gam_model)
print(k_diag)
```

**Plots of smooth effects**

```{r, fig.width=6, fig.height=4}
if (!is.null(params$smooth_plots)) {
  for (p in params$smooth_plots) {
    print(p)
  }
} else {
  cat("Smooth plots not provided.")
}
```

# Diagnostic Plots

**Manual Raw Residuals and Observed vs Predicted Plots**

```{r, fig.width=6, fig.height=4}
params$residuals_plot
params$obs_vs_pred_plot
```

**GAM Diagnostics (gratia::appraise)**

```{r}
if (!is.null(params$appraise_plot)) {
  print(params$appraise_plot)
} else {
  cat("Appraise plot not provided.")
}
```

# Multicollinearity Diagnostics

**GVIF Diagnostics**

```{r}
if (!is.null(params$gvif_tbl)) {
  params$gvif_tbl %>%
    dplyr::mutate(
      GVIF        = round(GVIF, 2),
      `GVIF_adj`  = round(`GVIF_adj`, 2)
    ) %>%
    kable(
      booktabs = TRUE,
      caption  = "GVIF and Adjusted GVIF for Parametric Terms"
    ) %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "condensed")
    )
} else {
  cat("GVIF not computed for this target (no subcategory collapsing applied).")
}
```

**Concurvity / VIF for Smooth Terms**

Note: Adjusted GVIFs account for the number of parameters in each term, making them comparable across terms with different degrees of freedom.   
For smooth terms, the worst-case concurvity shows the maximum dependence of one smooth on the others.

```{r}
has_re_smooth <- FALSE

if (!has_re_smooth) {
  conc_matrix <- mgcv::concurvity(params$gam_model, full = FALSE)$worst

  conc_matrix %>%
    as.data.frame() %>%
    tibble::rownames_to_column("Term") %>%
    dplyr::mutate(
      dplyr::across(
        -Term,
        ~ formatC(.x, format = "e", digits = 2)
      )
    ) %>%
    kable(
      booktabs = TRUE,
      caption  = "Worst-case concurvity diagnostics"
    ) %>%
    kable_styling(
      full_width = FALSE,
      position = "left",
      bootstrap_options = c("striped", "condensed")
    )

} else {
  cat("Concurvity diagnostics skipped: model contains random-effect smooths (bs = 're').")
}
```